---
title: "Investigating Missingness"
author: "Chris LeBoa"
date: "3/10/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Investigating Missingness 

```{r, include=FALSE}
# Libraries
library(tidyverse)
library(haven)

# Parameters
data_location <- "/Users/ChrisLeBoa/Dropbox/COVID_aerosols/clean data/2 25 21/mergeddata_22521.dta"
merged_data <- read_dta(file = data_location)


#Variables of intrest
list <-
  c(
  "roomvol",
  "co2average", 
  "tempstart",
  "humiditystart",
  "numcovidstart",
  "numpeopleavg", 
  "numstaffstart",
  "popdensityavg", 
  "totfan", 
  "numacon", 
  "numdoortotal", 
  "doortofloorarea",
  "numdoorclosed", 
  "numwintotal", 
  "numwinclosed",
  "wintofloorarea",
  "ach",
  "ventrateavg"
  )
```

```{r}
merged_data %>%
  mutate(across(list, is.na)) %>%  # replace all NA with TRUE and else FALSE
  pivot_longer(list, names_to = "var") %>%  # pivot longer
  filter(value) %>%   # remove the FALSE rows
  group_by(sampleid) %>%    # group by the ID
  summarise(`Missing Variables` = toString(var)) %>% 
  write_csv(here::here("output/missingness.csv"))# convert the variable names to a string column

```

#Window Open Area
```{r}
merged_data %>% 
  filter(is.na(openwinarea))
```

For the most part this is NA for when the number of windows open is 0. In four instances there Is also an NA for NumWinOpen 

We should discuss those four 

```{r}
merged_data %>% 
  filter(is.na(numwinopen)) %>% 
  select(sampleid, colldate, locationtype)
```

#Door Open Area 
For the total door open area these should all be 0 since they all have 0 doors open 

for icd1_r1 there are a list of 3 open doors but the open door area parameter reads 0. Why? 
```{r}
merged_data %>% 
  filter(is.na(opendoorarea)) %>% 
  select(numdooropen)

merged_data %>% 
  filter(is.na(smallestopendoorarea)) %>% 
  select(sampleid, opendoorarea, numdooropen)
```

For the two that are missing temp end measurements we should add the temp that they started at 
```{r}
merged_data %>% 
  filter(is.na(tempend)) %>% 
  select(sampleid, tempstart)
```


and for the sample missing a starting temperature we should add the ending temperature 
```{r}
merged_data %>% 
  filter(is.na(tempstart)) %>% 
  select(sampleid, tempend)
```

For the window to floor area measure we should hand calculate sample 5, add in 0 to the open window area for rooms with no open windows 
```{r}
merged_data %>% 
  filter(is.na(windoortofloorarea)) %>% 
  select(sampleid, openwinarea, numwintotal, numwinopen, floorarea)
```

For sample missing a humidity at the start we add the humidity from the end and vice versa
```{r}
merged_data %>% 
  filter(is.na(humiditystart)) %>% 
  select(sampleid, humidityend)
```

For the smallest open door area only the first sample had problems - says 3 doors but no area measure 

Rest should be 0 
```{r}
merged_data %>% 
  filter(is.na(smallestopendoorarea)) %>% 
  select(sampleid, numdooropen, numdoorfullopen, numdoortotal)
```



There are no missing middle ventilation rates so we do not have to worry about replacement 
```{r}
merged_data %>% 
  filter(is.na(ventratemid)) %>% 
  select(sampleid, ventratestart, ventrateend, ventrateavg)

merged_data %>% 
  filter(is.na(ventrateavg)) %>% 
  select(sampleid, ventratestart, ventrateend, ventratemid)
```

