---
output:
  pdf_document: default
  html_document: default
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(dplyr)
library(ggplot2)
library(reshape2)
library(haven)
library(knitr)
```

```{r data import}
#import data
d <- read_dta(file = "~/Dropbox/COVID aerosols/clean data/1 20 21/mergeddata_12021.dta")

#filtering to test data only
test <- d %>% filter(colldate != "")

test <- test %>%
  mutate(newwindow = ifelse(nearwindowopen %in% c("Closed", "N/A"), "Closed", "Open"),
         anyopen = ifelse(nearwindowopen == "Open" | neardooropen == "Open", "Either Open", "Both Closed"))
```

## CO2 averages by hospital, type of sampling space, and ventilation measures
```{r}
sum1 <- test %>%
  group_by(hosp) %>%
  summarise(n = n(), Mean = round(mean(co2average, na.rm = TRUE), digits = 1), SD = round(sd(co2average, na.rm = TRUE), digits = 1))

sum2 <- test %>%
  group_by(loctype) %>%
  summarise(n = n(), Mean = round(mean(co2average, na.rm = TRUE), digits = 1), SD = round(sd(co2average, na.rm = TRUE), digits = 1))

sum3 <- test %>%
  group_by(nearwindowopen) %>%
  summarise(n = n(), Mean = round(mean(co2average, na.rm = TRUE), digits = 1), SD = round(sd(co2average, na.rm = TRUE), digits = 1))

sum4 <- test %>%
  group_by(neardooropen) %>%
  summarise(n = n(), Mean = round(mean(co2average, na.rm = TRUE), digits = 1), SD = round(sd(co2average, na.rm = TRUE), digits = 1))

colnames(sum1) <- c("Hospital", "n", "Mean", "SD")
colnames(sum2) <- c("Type of sampling space", "n", "Mean", "SD")
colnames(sum3) <- c("Nearest window open", "n", "Mean", "SD")
colnames(sum4) <- c("Nearest door open", "n", "Mean", "SD")

kable(sum1)
kable(sum2)
kable(sum3)
kable(sum4)

```

\newpage
## CO2 ppm over time during data collection 
```{r}
co2.data <- test %>% select(sampleid, co2startnew, co2_5minnew, co2_10minnew,
                        co2_15minnew, co2_20minnew, co2_25minnew, co2endnew)

co2.long <- melt(co2.data, value.name = "co2value")

co2.long <- co2.long %>%
  mutate(min = ifelse(variable == "co2startnew", 0, 
               ifelse(variable == "co2_5minnew", 5,
               ifelse(variable == "co2_10minnew", 10,
               ifelse(variable == "co2_15minnew", 15,
               ifelse(variable == "co2_20minnew", 20,
               ifelse(variable == "co2_25minnew", 25,
               ifelse(variable == "co2endnew", 30, NA))))))))

ggplot(data = co2.long, aes(x = min, y = co2value, group = sampleid, color = sampleid)) +
  geom_line() + 
  theme(legend.position = "none") +
  labs(y = "CO2 (ppm)", x= "Minute of data collection")
```
\newpage
**Average CO2 value of 30-minute air sampling period by qPCR result**  
```{r}
results <- test %>%
  filter(sampletype == "TEST")

ggplot(data = results, aes(x = result, y = co2average)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(x = result, y = co2average, color = loctype)) +
  labs(x = "qPCR result", y = "Average CO2 (ppm)", color = "Type of\nsampling space")
```

\newpage  
  
```{r}
#create long dataset for 0, 15, 30 min mark
peop.data <- test %>% select(sampleid, numpeoplestart, numpeoplemid, numpeopleend)

peop.long <- melt(peop.data, value.name = "numpeople")

peop.long <- peop.long %>%
mutate(min = ifelse(variable == "numpeoplestart", 0, 
               ifelse(variable == "numpeoplemid", 15,
               ifelse(variable == "numpeopleend", 30, NA))))

popdens.data <- test %>% select(sampleid, popdensitystart, popdensitymid, popdensityend)

popdens.long <- melt(popdens.data, value.name = "popdensity")

popdens.long <- popdens.long %>%
mutate(min = ifelse(variable == "popdensitystart", 0, 
               ifelse(variable == "popdensitymid", 15,
               ifelse(variable == "popdensityend", 30, NA))))

total.long <- left_join(peop.long, co2.long, by = c("sampleid", "min"))
total.long <- merge(total.long, popdens.long, by = c("sampleid", "min"))
total.long <- total.long %>% select(-variable, -variable.x, -variable.y)

test2 <- test %>% 
  select(sampleid, roomvol, loctype, newwindow, nearwindowopen, numwinopen, openwinarea, neardooropen, numdooropen, opendoorarea, anyopen) %>%
  mutate(openwindoorarea = openwinarea + opendoorarea)

total.long <- merge(total.long, test2, by = "sampleid")
```
## Number of people in the room vs CO2 values, by whether the closest* window was open  
```{r}
ggplot(data = total.long, aes(x = numpeople, y = co2value, color = loctype)) + 
  geom_point() +
  geom_line(aes(group = sampleid)) +
  facet_wrap(~newwindow) +
  labs(y = "CO2 (ppm)", x= "Number of people in sampling space", color = "Type of\nsampling space")
```
*Rooms where there were no windows were categorized as "closed"  

## Number of people in the room vs CO2 values, by whether the closest door was open
```{r}
ggplot(data = total.long, aes(x = numpeople, y = co2value, color = loctype)) + 
  geom_point() +
    geom_line(aes(group = sampleid)) +
  facet_wrap(~neardooropen) +
  labs(y = "CO2 (ppm)", x= "Number of people in sampling space", color = "Type of\nsampling space")
```

## Number of people in the room vs CO2 values, by whether the closest door or window* was open  
```{r}
ggplot(data = total.long, aes(x = numpeople, y = co2value, color = loctype)) + 
  geom_point() +
  geom_line(aes(group = sampleid)) +
  facet_wrap(~anyopen) +
  labs(y = "CO2 (ppm)", x= "Number of people in sampling space", color = "Type of\nsampling space")
```
*Rooms where there were no windows were categorized as closed  
\newpage
  
## CO2 ppm by room area and number of people in the sampling space
```{r}
ggplot(data = total.long, aes(x = roomvol, y = numpeople, color = co2value)) +
  geom_point() +
  labs(y = "Number of people in sampling space", x= "Room volume (ft cub)", color = "CO2 (ppm)")
```
\newpage
  
## CO2 ppm by density of people in the sampling space
```{r}
ggplot(data = total.long, aes(x = popdensity, y = co2value)) +
  geom_point() +
  geom_smooth(method = "gam") + 
  labs(y = "CO2 (ppm)", x= "People per cubic foot")
```
\newpage
  
## CO2 ppm by room density and open window and door area
```{r}
ggplot(data = total.long, aes(x = popdensity, y = openwindoorarea, color = co2value)) +
  geom_point() +
  labs(x = "People per cubic foot", y= "Total area of open windows and doors", color = "CO2 (ppm)") +
  xlim(0,0.01)
```

```{r}
ggplot(data = total.long) +
  geom_jitter(aes(y = co2value, x = loctype)) + 
  geom_hline(yintercept = 500, color = "red") +
  labs(y = "CO2 value (ppm)", x = "Location Type")
```



