---
editor_options:
  chunk_output_type: inline
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r message=FALSE, warning=FALSE}
#import data
source(here::here("configuration.R"))

#library(SmartEDA)
library(reshape2)
library(GGally)
library(knitr)
library(dummies)
library(glmnet)
library(plotmo)
```

```{r}
d <- read_dta(paste0(dropboxDir, "/clean data/Final dataset/mergeddata_final.dta"))
d[d == ""] <- NA

d <- d %>%
  filter(!is.na(Q)) %>%
  mutate(ventratelog10 = log10(ventrateavg),
         Qlog10 = log10(Q),
         windowordooropennear = ifelse(windowopennear == 1 | dooropennear == 1, 1, 0),
         windowsopp = as.numeric(windowsopp))

#exploratory data analysis
#ExpCatViz(d)
#ExpNumViz(d)

#select ventilation measures of interest
vent.measures <- d %>%
  select(sampleid, roomheight, roomvol, surfacearea, wallarea, floorarea, 
         openwinarea, wintofloorarea, opendoorarea, doortofloorarea, 
         totalopenarea, windoortofloorarea, windoortovolume, 
         acon, fanson, tempstart, humiditystart, windowordooropennear, windowsopp)

```

### Visualize correlation of covariates
```{r}
#test correlation of measures for numeric variables
ggcorr(vent.measures[,-c(1)], label = TRUE, label_round = 2, label_size = 3, hjust = 1)

vent.measures <- vent.measures %>% select(-surfacearea, -wallarea, -windoortovolume, -wintofloorarea, -doortofloorarea)
```

All are not positively correlated, which is good. Can gain information from variability. Remove 
* Surface area and wall area as they are almost perfectly correlated with floor area and room volume
* Window + door area to room volume ratio as it is almost perfectly correlated with window + door area to floor area ratio
* Window area to floor area ratio and door area to floor area ratio as there will be more utility in having a combined measure

### Look at bivariate relationships
```{r}
d$locationtype <- factor(d$locationtype, levels = c("Private room", "Open ward", "ICU", "OPD", "Other"))

ggplot(data = d, aes(x = roomheight, y = Qlog10)) + 
  geom_smooth(method = 'lm', ) +
  geom_point(aes(color = locationtype)) 

#can see clustering by location type - check facet by location type
ggplot(data = d, aes(x = roomheight, y = Qlog10)) + 
  geom_smooth(method = 'lm', ) +
  geom_point(aes(color = locationtype)) +
  facet_wrap(~locationtype)

#check sparsity for binary variables
ggplot(data = d, aes(x = as.factor(windowsopp), y = Qlog10)) +
 geom_boxplot() +
 geom_jitter() +
 facet_wrap(~locationtype)

```
Decision: include location type as fixed effect, however there is sparsity

### Check distributions for outliers
```{r}
plots <- d %>% select(all_of(names(vent.measures)))
plots <- melt(plots)
Qlog <- d %>% select(sampleid, Qlog10)
plots <- left_join(plots, Qlog, by = "sampleid")
ggplot(data = plots) + geom_point(aes(x = value, y = Qlog10)) + facet_wrap(~variable, scales = "free")
```

### Univariable models using linear regression
```{r}
#filter out X variable outliers for univariate models (need to use separate dataset for each model to maximize sample size)
d1 <- d %>% filter(opendoorarea < 30)
d2 <- d %>% filter(windoortofloorarea < 0.8)
d3 <- d %>% filter(roomvol < 4000)

#univariate linear models
names <- c(names(vent.measures[-1]), "numpeopleavg", "locationtype")
all.res <- NULL
residuals <- NULL
#loop through using subsetted data for each
for (i in 1:length(names)){
  if(names[[i]] %in% c("opendoorarea")){
      model <- lm(as.formula(paste("Qlog10 ~ ", names[[i]], sep = "")), 
                  data = d1)
      res <- broom::tidy(model, conf.int = TRUE)
      res <- cbind(res[2,1], res[2,c(2,6,7,5)])
      all.res <- as.data.frame(rbind(all.res, res))
      resids <- cbind(term = res$term, fitted = predict(model), residuals = resid(model))
      residuals <- rbind(residuals, resids)
  } else if (names[[i]] %in% c("windoortofloorarea")){
      model <- lm(as.formula(paste("Qlog10 ~ ", names[[i]], sep = "")), 
                  data = d2)
      res <- broom::tidy(model, conf.int = TRUE)
      res <- cbind(res[2,1], res[2,c(2,6,7,5)])
      all.res <- as.data.frame(rbind(all.res, res))
      resids <- cbind(term = res$term, fitted = predict(model), residuals = resid(model))
      residuals <- rbind(residuals, resids)
  } else if (names[[i]] %in% c("roomvol", "floorarea")){
      model <- lm(as.formula(paste("Qlog10 ~ ", names[[i]], sep = "")), 
                  data = d3)
      res <- broom::tidy(model, conf.int = TRUE)
      res <- cbind(res[2,1], res[2,c(2,6,7,5)])
      all.res <- as.data.frame(rbind(all.res, res))
      resids <- cbind(term = res$term, fitted = predict(model), residuals = resid(model))
      residuals <- rbind(residuals, resids)
  } else if (names[[i]] %in% c("locationtype")){
      model <- lm(as.formula(paste("Qlog10 ~ ", names[[i]], sep = "")), 
                  data = d)
      res <- broom::tidy(model, conf.int = TRUE)
      res <- cbind(res[c(2:5),1], res[c(2:5),c(2,6,7,5)])
      all.res <- as.data.frame(rbind(all.res, res))
      resids <- cbind(term = res$term, fitted = predict(model), residuals = resid(model))
      residuals <- rbind(residuals, resids)
  } else {
      model <- lm(as.formula(paste("Qlog10 ~ ", names[[i]], sep = "")), 
                  data = d)
      res <- broom::tidy(model, conf.int = TRUE)
      res <- cbind(res[2,1], res[2,c(2,6,7,5)])
      all.res <- as.data.frame(rbind(all.res, res))
      resids <- cbind(term = res$term, fitted = predict(model), residuals = resid(model))
      residuals <- rbind(residuals, resids)
    }
}

#check residuals
residuals <- as.data.frame(residuals)
ggplot(data = residuals, aes(x = as.numeric(fitted), y = as.numeric(residuals))) + 
  geom_point() +
  geom_smooth(method = 'gam') +
  facet_wrap(~term)
```
Decision: include temperature, humidity, window open nearby, and door open nearby in LASSO regression as a nuisance parameters

### Multivariable model using LASSO regression
```{r}
#multivariate model
#check missingness - need complete cases
check <- colSums(is.na(vent.measures))

#filter variables with missing data and extreme outliers
d.new <- d %>% 
  filter(opendoorarea < 30 & windoortofloorarea < 0.8, roomvol < 4000) %>%
  dplyr::select(all_of(names(vent.measures)), numpeopleavg, Qlog10, locationtype,
                -distwindow, -distdoor, -sampleid) %>%
  mutate(acon = as.numeric(acon),
         fanson = as.numeric(fanson),
         windowsopp = as.numeric(windowsopp))
  

#create matrix of indicator variables for location type
loctype.dummy <- dummy(d.new$locationtype)
colnames(loctype.dummy) <- c("Private room", "Open ward", "ICU", "OPD", "Other")

#bind together. make private room the reference category
vent.matrix <- as.matrix(cbind(d.new[, c(1:(ncol(d.new)-2))], loctype.dummy[,-1]))

#run LASSO regression
set.seed(123)
mod <- glmnet::glmnet(vent.matrix, d.new$Qlog10, family = "gaussian", alpha = 1)
cvfit <- glmnet::cv.glmnet(vent.matrix, d.new$Qlog10, family = "gaussian", 
                           alpha = 1, type.measure = "mse", nfolds = 5)

#extract coefficients
glmcoef <- coef(mod, cvfit$lambda.min)
coefficient <- data.frame(glmcoef[-1,1])
coefficient <- cbind(rownames(coefficient), coefficient)
rownames(coefficient) <- c(1:nrow(coefficient))
colnames(coefficient) <- c("Variable", "Coefficient")
coefficient  %>% arrange(-Coefficient) %>% kable()

#visualize to check LASSO shrinkage
coef.increase <- coefficient[which(coefficient$Coefficient>0), 1]
coef.decrease <- coefficient[which(coefficient$Coefficient<0), 1]

allnames<-names(coef(mod)[,
    ncol(coef(mod))][order(coef(mod)[,
    ncol(coef(mod))],decreasing=TRUE)])

allnames<-setdiff(allnames,allnames[grep("Intercept",allnames)])

cols<-rep("gray",length(allnames))
cols[allnames %in% coef.increase]<-"green"      # higher ventilation is good
cols[allnames %in% coef.decrease]<-"red"        # lower ventilation is not
plot_glmnet(mod,label=TRUE,s=cvfit$lambda.min,col=cols)
```

### Multivariable models
```{r}
#extract nonzero coefficients to put in multivariable model
nonzero <- coefficient$Variable[which(coefficient$Coefficient != 0)]
nonzero

model3 <- lm(Qlog10 ~ roomheight + floorarea + opendoorarea + totalopenarea + acon + fanson +
               tempstart + humiditystart + windowordooropennear + windowsopp + 
               numpeopleavg + locationtype, data = d.new)

#extract coefficients and confidence intervals
adj <- broom::tidy(model3, conf.int = TRUE)
adj <- adj %>% filter(term != "(Intercept)") %>% select(term, estimate, p.value, conf.low, conf.high) %>%
  rename(Variable = term)
```

Coefficients of 0 are less important - have been shrunk to 0 using lambda penalization  

### Obtain elasticity from unadjusted and adjusted coefficients
```{r}
#unadjusted measures
unadj <- all.res %>% rename(Variable = term)
unadj$Variable <- str_remove(unadj$Variable, "locationtype")

#lasso coefficient
fullmodels <- left_join(unadj, coefficient, by = "Variable")

#adjusted measures
adj$Variable <- str_remove(adj$Variable, "locationtype")
fullmodels <- left_join(fullmodels, adj, by = "Variable")

for (i in 1:nrow(fullmodels)){
  #no elasticity for location type coefficients
  if(fullmodels[i,1] %in% c("Open ward", "ICU", "OPD", "Other")){
    fullmodels$adj.elas[i] <- NA
    fullmodels$adj.elas.lower[i] <- NA
    fullmodels$adj.elas.upper[i] <- NA
  } else{
    var <- fullmodels$Variable[i]
    temp <- as.matrix(d.new[,var])
    fullmodels$adj.elas[i] <- signif(fullmodels$estimate.y[i] * ((mean(temp, na.rm = TRUE) / mean(d.new$Qlog10))), 2)
    fullmodels$adj.elas.lower[i] <- signif(fullmodels$conf.low.y[i] * ((mean(temp, na.rm = TRUE) / mean(d.new$Qlog10))), 2)
    fullmodels$adj.elas.upper[i] <- signif(fullmodels$conf.high.y[i] * ((mean(temp, na.rm = TRUE) / mean(d.new$Qlog10))), 2)
  }
}

labels <- c("Room height (m)", "Room volume (m^3)", "Floor area (m^2)",  
            "Open window area (m^2)", "Open door area (m^2)", 
            "Total open window and door area (m^2)",  
            "Ratio of open window and door area to floor area",
             "AC on",  "Fans on", "Temperature (C)", "Humidity",
            "Open window or door nearby", "Cross ventilation present",
            "Average number of people", "Open Ward", "ICU", "OPD", "Other")

fullmodels <- cbind(labels = labels, fullmodels)

fullmodels <- fullmodels %>% filter(Variable != "tempstart" & Variable != "humiditystart" &
                                      Variable != "windowordooropennear")

table <- fullmodels %>%
  mutate(unadjusted = paste(signif(estimate.x, 2), " (", signif(conf.low.x, 2), ", ", 
                            signif(conf.high.x, 2), ")", sep = ""),
         adjusted = paste(signif(estimate.y, 2), " (", signif(conf.low.y, 2), ", ", 
                          signif(conf.high.y, 2), ")", sep = ""),
         elasticity = paste(signif(adj.elas, 2), " (", signif(adj.elas.lower, 2), ", ", 
                          signif(adj.elas.upper, 2), ")", sep = "")) %>%
  select(labels, unadjusted, Coefficient, adjusted, elasticity) %>%
  mutate(Coefficient = signif(Coefficient, 2))

table$adjusted[table$adjusted == "NA (NA, NA)"] <- "-"
table$elasticity[table$elasticity == "NA (NA, NA)"] <- "-"

colnames(table) <- c("Variable", "Unadjusted mean difference (95% CI)", "LASSO coefficient", "Adjusted mean difference (95% CI)1", "Elasticity2")

table %>% kable(align = "lcccc")
```
1 Adjusted for temperature, humidity, and whether a window or door was open nearby as nuisance parameters
2 Interpreted as the percentage change in ventilation rate for a percent increase in the given variable

```{r}
#visualize elasticity
elasticity <- fullmodels %>%
  select(labels, adj.elas, adj.elas.lower, adj.elas.upper)

sig.names <- na.omit(elasticity[elasticity$adj.elas.lower > 0, 1])

elasticity.long <- melt(elasticity, id = "labels")

elasticity.long <- elasticity.long %>%
  filter(!is.na(value)) %>%
  mutate(sig = as.factor(ifelse(labels %in% sig.names, "Significant", "Not Significant")))

plot <- ggplot(data = elasticity.long) +
  geom_point(aes(x = reorder(labels, value), y = value, color = sig))+
  geom_line(aes(x = labels, y = value, group = labels, color = sig)) +
  geom_hline(yintercept = 0, colour = "red") +
  labs(color = "Significance", y = "Percent change in absolute ventilation (log10 L/s) \n for a 1% change in parameter", x = "Variable") +
  theme_minimal() +
  scale_x_discrete("", labels = c( "Fans on",
                                   expression(paste("Open door area (m"^2,")")),
                                   "Cross ventilation present",
                                   "Average number of people",
                                   "AC on",
                                   expression(paste("Floor area (m"^2,")")),
                                   expression(paste("Total open window and door area (m"^2,")")),
                                   "Ceiling height")) +
  coord_flip() +
  scale_color_manual(values = c("dark gray", "black")) +
  theme(legend.text = element_text(size = 9),
        legend.title = element_blank())

ggsave(filename = "~/Documents/COVID/elasticity.jpg",
       width = 10, height = 6)
```
