---
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r message=FALSE, warning=FALSE}
#import data
rm(list=ls())
source(here::here("configuration.R"))

#library(SmartEDA)
library(reshape2)
library(GGally)
library(knitr)
library(dummies)
library(glmnet)
library(plotmo)
```

```{r}
d <- read_dta(paste0(dropboxDir, "/clean data/Final dataset/mergeddata_final.dta"))
d[d == ""] <- NA

#reframe these measures to be - is a window/door open nearby? 
d <- d %>%
  mutate(windowopennear = ifelse(nearwindowopen == "Open", 1, 
                                 ifelse(nearwindowopen == "Closed", 0, NA)),
         windowopennear = ifelse(is.na(windowopennear), 0, windowopennear),
         dooropennear = ifelse(neardooropen == "Open", 1, 
                                 ifelse(neardooropen == "Closed", 0, NA)),
         dooropennear = ifelse(is.na(dooropennear), 0, dooropennear))

#exploratory data analysis
#ExpCatViz(d)
#ExpNumViz(d)

#select ventilation measures of interest
vent.measures <- d %>%
  select(sampleid, roomheight, roomvol, surfacearea, wallarea, floorarea, 
         openwinarea, wintofloorarea,
         opendoorarea, doortofloorarea, 
         totalopenarea, windoortofloorarea,
         numfanon, numacon, tempstart, humiditystart,
         distwindow, distdoor, windowopennear, dooropennear)

```

### Visualize correlation of covariates
```{r}
#test correlation of measures for numeric variables
ggcorr(vent.measures[,-c(1)], label = TRUE, label_round = 2, label_size = 3, hjust = 1)
```

All are not positively correlated, which is good. Can gain information from variability. Remove surface area and wallarea as they are almost perfectly correlated with room volume.

### Look at bivariate relationships
```{r}
vent.measures <- vent.measures %>% select(-surfacearea, -wallarea)
d$locationtype <- factor(d$locationtype, levels = c("Open ward", "Bathroom", "Cabin", "Canteen", "Doffing room", "Sample collection", "OPD"))

ggplot(data = d, aes(x = openwinarea, y = ventrateavg)) + 
  geom_smooth(method = 'gam', ) +
  geom_point(aes(color = locationtype)) 

#can see clustering by location type - check facet by location type
ggplot(data = d, aes(x = openwinarea, y = ventrateavg)) + 
  geom_smooth(method = 'gam', ) +
  geom_point(aes(color = locationtype)) +
  facet_wrap(~locationtype)

#remove canteen for data sparsity
total.filt <- d %>% filter(locationtype != "Canteen")

#filter out X variable outliers for univariate models (need to use separate dataset for each model to maximize sample size)
total.filt1 <- total.filt %>% filter(roomvol < 5000) #for all room measurements
total.filt2 <- total.filt %>% filter(wintofloorarea < 0.25)
total.filt3 <- total.filt %>% filter(opendoorarea < 30)
total.filt4 <- total.filt %>% filter(windoortofloorarea < 0.2)
```
Decision: include location type as a fixed effect in all models

### Univariable models using linear regression models
```{r}
#univariate linear models
names <- names(vent.measures[-1])
all.res <- NULL

for (i in 1:length(names)){
  if(names[[i]] %in% c("roomvol", "floorarea")){
      model <- lm(as.formula(paste("ventrateavg ~ ", names[[i]], " + locationtype", sep = "")), 
                  data = total.filt1)
      res <- broom::tidy(model, conf.int = TRUE)
      res <- cbind(res[2,1], res[2,c(2,6,7,5)])
      all.res <- as.data.frame(rbind(all.res, res))
  } else if (names[[i]] %in% c("wintofloorarea")){
      model <- lm(as.formula(paste("ventrateavg ~ ", names[[i]], " + locationtype", sep = "")), 
                  data = total.filt2)
      res <- broom::tidy(model, conf.int = TRUE)
      res <- cbind(res[2,1], res[2,c(2,6,7,5)])
      all.res <- as.data.frame(rbind(all.res, res))
  } else if (names[[i]] %in% c("opendoorarea")){
      model <- lm(as.formula(paste("ventrateavg ~ ", names[[i]], " + locationtype", sep = "")), 
                  data = total.filt3)
      res <- broom::tidy(model, conf.int = TRUE)
      res <- cbind(res[2,1], res[2,c(2,6,7,5)])
      all.res <- as.data.frame(rbind(all.res, res))
  } else if (names[[i]] %in% c("windoortofloorarea")){
      model <- lm(as.formula(paste("ventrateavg ~ ", names[[i]], " + locationtype", sep = "")), 
                  data = total.filt4)
      res <- broom::tidy(model, conf.int = TRUE)
      res <- cbind(res[2,1], res[2,c(2,6,7,5)])
      all.res <- as.data.frame(rbind(all.res, res))
  } else {
      model <- lm(as.formula(paste("ventrateavg ~ ", names[[i]], " + locationtype", sep = "")), 
                  data = total.filt)
      res <- broom::tidy(model, conf.int = TRUE)
      res <- cbind(res[2,1], res[2,c(2,6,7,5)])
      all.res <- as.data.frame(rbind(all.res, res))
    }
}

#visualize univariate coefficients
sig.names <- all.res[all.res$p.value < 0.05,1]

all.res.long <- melt(all.res[,-5], id = "term")
all.res.long <- all.res.long %>%
  rename(Variable = term) %>%
  mutate(sig = as.factor(ifelse(Variable %in% sig.names, "Significant", "Not Significant")))

ggplot(data = all.res.long) +
  geom_point(aes(x = reorder(Variable, value), y = value, color = sig))+
  geom_line(aes(x = Variable, y = value, group = Variable, color = sig)) +
  geom_hline(yintercept = 0, colour = "red") +
  labs(color = "Significance", y = "Mean Difference for a 1-unit change", x = "Variable") +
  coord_flip() +
  scale_color_manual(values = c("dark gray", "black")) 

```


Decision: include temperature, humidity, window open nearby, and door open nearby in multivariable model as a nuisance parameter  

### Multivariable model using LASSO regression
```{r}
#multivariate model
#check missingness - need complete cases
check <- colSums(is.na(vent.measures))

#filter variables with missing data and extreme outliers
vent <- total.filt %>% 
  dplyr::select(all_of(names(vent.measures)), locationtype, ventrateavg, 
                -distwindow, -distdoor, -sampleid) %>% 
  filter(roomvol < 5000 & wintofloorarea < 0.25 & 
           opendoorarea < 30 & windoortofloorarea < 0.2)

#create matrix of predictor variables
loctype.dummy <- dummy(vent$locationtype)
colnames(loctype.dummy) <- c("Open Ward", "Bathroom", "Cabin", "Doffing Room", "Sample Collection", "OPD")
vent.matrix <- as.matrix(cbind(vent[complete.cases(vent), 1:15], loctype.dummy[,-1]))

#run LASSO regression
set.seed(123)
mod <- glmnet::glmnet(vent.matrix, vent$ventrateavg, family = "gaussian", alpha = 1)
cvfit <- glmnet::cv.glmnet(vent.matrix, vent$ventrateavg, family = "gaussian", alpha = 1, type.measure = "mse")

#extract coefficients
glmcoef <- coef(mod, cvfit$lambda.min)

coef.increase<-dimnames(glmcoef[glmcoef[,1]>0,0])[[1]]
coef.decrease<-dimnames(glmcoef[glmcoef[,1]<0,0])[[1]]

#get ordered list of variables as they appear at smallest lambda
allnames<-names(coef(mod)[,
    ncol(coef(mod))][order(coef(mod)[,
    ncol(coef(mod))],decreasing=TRUE)])

#remove intercept
allnames<-setdiff(allnames,allnames[grep("Intercept",allnames)])

#assign colors
cols<-rep("gray",length(allnames))
cols[allnames %in% coef.increase]<-"green"      # higher ventilation is good
cols[allnames %in% coef.decrease]<-"red"        # lower ventilation is not

```

### Plot shrinkage
```{r}
plot_glmnet(mod,label=TRUE,s=cvfit$lambda.min,col=cols)

#get coefficient table
coefficient <- data.frame(glmcoef[-1,1])
            
coefficient <- cbind(rownames(coefficient), coefficient)
rownames(coefficient) <- c(1:nrow(coefficient))
colnames(coefficient) <- c("Variable", "Coefficient")

labels <- c("Room height (m)", "Room volume (m^3)","Floor area (m^2)",  "Open window area (m^2)",
            "Ratio of open window area to floor area",
            "Open door area (m^2)", "Ratio of open door area to floor area",
            "Total open window and door area (m^2)",  "Ratio of open window and door area to floor area",
             "Number of fans on",  "Number of AC on", "Temperature (C)", "Humidity",
            "Open window nearby", "Open door nearby", "Bathroom", "Cabin", "Doffing Room", 
            "Sample Collection", "OPD")

coefficient %>% mutate(Variable = labels) %>% arrange(-Coefficient) %>% kable()

```
Reference level for location type = "Open Ward"  
Coefficients of 0 are less important - have been shrunk to 0 using lambda penalization  

### Obtain elasticity from unadjusted and adjusted coefficients
```{r}
coefficient <- coefficient %>%
  filter(Coefficient != 0) %>%
  filter(Variable !=  "Bathroom" & Variable != "Cabin"
         & Variable != "Canteen" & Variable !=  "Doffing Room" & Variable !=  "Sample Collection"
         & Variable != "OPD") 

#elasticity
all.res <- all.res %>% rename(Variable = term)
fullmodels <- left_join(all.res, coefficient, by = "Variable")
fullmodels <- fullmodels %>% filter(Variable != "distwindow" & Variable != "distdoor")

for (i in 1:nrow(fullmodels)){
  var <- fullmodels$Variable[i]
  temp <- as.matrix(vent[,var])
  fullmodels$unadj.elas[i] <- round(fullmodels$estimate[i] * (mean(temp, na.rm = TRUE) / mean(vent$ventrateavg)), 3)
  fullmodels$adj.elas[i] <- round(fullmodels$Coefficient[i] * (mean(temp, na.rm = TRUE) / mean(vent$ventrateavg)), 3)
}

fullmodels <- cbind(labels = labels[-c(16:20)], fullmodels)

fullmodels <- fullmodels %>%
  mutate(unadjusted = paste(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")", sep = ""),
         lasso = round(Coefficient, 2), Variable = labels) %>%
  select(Variable, unadjusted, unadj.elas, lasso, adj.elas) %>%
  arrange(-adj.elas) 

fullmodels[is.na(fullmodels)] <- "-"
colnames(fullmodels) <- c("Variable", "Mean difference (95% CI)*", "Elasticity^ from unadjusted model", "LASSO coefficient", "Elasticity from adjusted model")

fullmodels %>% kable()
```
*Estimated using linear regression models adjusted for location type  
^Interpreted as the percentage change in ventilation rate for a percent increase in the given variable
